---
title: "Modeling NRG"
author: "Adam"
date: "March 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The Datafiles
The data are stored in .csv files containing the eye, head and gaze positions in ms, and whether there was also a spike in that millisecond (rasters). It has the subject letter, the trial number (which includes the date and monkey letter), as well as the type of trial: ramp, two-trajectories, gap, delayed gaze shifts, etc. There is more data in the original H5 files that the .csv files were extracted from. The name of the h5 file is the same as the trial number. eg: UB01may15_04, is trial number 4 from monkey U collected on may 1st 2015. The second letter indicates which of several files may have been recorded during the session. UA01may15, UB01may15, etc.

```{r load.libraries, message=FALSE}
library(checkpoint)
checkpoint('2016-03-20')
install.packages('relaimpo')
install.packages('leaps')
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
# library(broom)
# library(grid)
library(relaimpo)
library(leaps)
#library(data.table)
library(stringr)

```

```{r helperfunctions}

spikedensity<-function (rasters,sd=100) {
  gsize<- sd*10
  g<-dnorm(-gsize:gsize,mean=0,sd=sd)
  sdf<-convolve(rasters,g,type="open")
  sdf<-sdf[gsize:(length(sdf)-(gsize+1))]*1000
  sdf
}

dynamiclead<-function(p,lags=seq(10,300,by=10),fit.eq='sdflag~head+hv+ha+eye+ev+ea') {

  rsq<-NULL
  for (i in 1:length(lags)) {
    if (lags[i] > 0){
      p$sdflag<-dplyr::lag(p$sdf,lags[i])
    }
    else{
      p$sdflag<-dplyr::lead(p$sdf,lags[i]*-1)
    }

    rsq[i]<- summary(lm(fit.eq,data=p))$r.squared
  }
  #return(rsq)
  return(lags[rsq==max(rsq)])
}

findSaccades<-function(ev,threshold=40){
  
  i<-which(abs(ev)>threshold) #find all the times when speed > threshold
  sacoff<-which(diff(i)>15) #minimum duration of an accepted saccade
  sacon<-c(1,sacoff+1) #first saccade
  sacoff<-c(sacoff,length(i)) #end of last saccade
  saccade.onset<-i[sacon] #get actual times
  saccade.offset<-i[sacoff] 
  return(data.frame(saccade.onset,saccade.offset))
}

markSaccades<-function(ev,buffer=15,threshold=40){
  #this function finds and marks saccades given a velocity input
  stimes<-findSaccades(ev,threshold)
  
  #remove saccades without enough data at the end of the file, based on buffer size
  toolong<- stimes$saccade.offset> length(ev)-buffer
  tooshort<- stimes$saccade.onset<buffer+1
  stimes<- filter(stimes, !tooshort, !toolong)
  
  nsaccades=nrow(stimes)

  stimes$saccade.onset=stimes$saccade.onset-buffer
  stimes$saccade.offset=stimes$saccade.offset+buffer
  
  s<-1:length(ev)*0
  
  for (k in 1:nsaccades){
    s[stimes$saccade.onset[k]:stimes$saccade.offset[k]]<- k
    if(k>1){
      s[stimes$saccade.offset[k-1]:stimes$saccade.onset[k]]<-(k*-1)
    }
  }
  s[1:stimes$saccade.onset[1]]<- -1
  s[stimes$saccade.offset[nrow(stimes)]:length(s)]<- (nrow(stimes)*-1)-1
  return(s)
}

parabolicdiff <- function(pos,n=7){
  q <- sum(2*((1:n)^2))
  convoutput<- convolve(pos,c(-n:-1, 1:n),type="open")
  convoutput<- convoutput[(n*2):(length(pos)-((n*2)+1))]
  vels<- c(array(convoutput[1],dim=n*2),convoutput,array(convoutput[length(convoutput)],dim=n*2))
  vels <- vels/q*1000
}

```
```{r load.data}
path<- "C:/Users/setup/Desktop/NeurophysNRG/data/"
savename<- 'allNeuronsLagSplit.RDS'

if (!file.exists(savename)) {
  message('Reading data from .csv')
  #get names of all files in path
  files <- list.files(path=path,pattern='*.csv')
  
  nfiles<-length(files)
  # nfiles=1
  if (nfiles>0){
    message(paste('Loading: ',files))
    loadedfiles <- lapply(paste(path,files,sep=''),read.csv)
  }else{
    stop('********NO NEW CELLS********')
  }
  
  t<-data.frame()
  buffer<- 200
  for (i in 1:nfiles) {
    f<- files[i]
    temp=loadedfiles[[i]]
    
    # leadtime<-dynamiclead(temp)
    temp %>%
      separate(trialnum,c('neuron','trialnum'),sep='trial_',remove=TRUE) %>%
      mutate(trialnum=as.numeric(trialnum)) %>%
      group_by(trialnum) %>%
      mutate(ev=parabolicdiff(eye),
             hv=parabolicdiff(head),
             gv=parabolicdiff(gaze),
             ea=parabolicdiff(ev),
             ha=parabolicdiff(hv),
             sdf=spikedensity(rasters,sd=10))->
      temp
    
    t <-rbind(t,temp)
  }
  saveRDS(t,savename)
} else{
  message('loading data from RDS')
  t<-readRDS(savename)
}

```

```{r split.lr}
# t %>% 
#   mutate(eyeL=replace(eye, eye>0, 0),
#          evL=replace(ev,ev>0,0),
#          eaL=replace(ea,ea>0,0),
#          headL=replace(head, head>0,0),
#          hvL=replace(hv,hv>0,0),
#          haL=replace(ha,ha>0,0),
#          
#          eye=replace(eye,eye<0,0),
#          ev=replace(ev,ev<0,0),
#          ea=replace(ea,ea<0,0),
#          head=replace(head,head<0,0),
#          hv=replace(hv,hv<0,0),
#          ha=replace(ha,ha<0,0)) ->
#   t
```

```{r shift.sdf}
# #This is a hack because dplyr doesn't allow the lag window function to take a different value for each group
# n<-unique(t$neuron)
# fit.eq='sdflag~eye+ev+ea+head+hv+ha+eyeL+evL+eaL+headL+hvL+haL'
# newt<-data.frame()
# for (i in 1:length(n)){
#   temp<-filter(t,neuron==n[[i]])
#   l<-dynamiclead(temp,lags=seq(10,150,by=10),fit.eq=fit.eq)
#   temp<-mutate(temp,leadtime=l,sdflag=lag(sdf,l))
#   newt<-rbind(newt,temp)
# }
# t<-newt
# newt<-NULL
# saveRDS(t,'allNeuronsLag.RDS')
```


```{r calc.relaimpo}

t %>% 
  group_by(neuron) %>% 
  do(b=calc.relimp(lm(sdflag~eye+ev+ea+head+hv+ha,data=.))) -> 
  rimp

r <- data.frame()
r2 <- NULL

for (i in 1:nrow(rimp)){
  plot(rimp$b[[i]], main=rimp$neuron[i])
  b<- rimp$b[[i]]$lmg
  r<-rbind(r,b)
  r2<- c(r2,rimp$b[[i]]$R2)
  
}

r$neuron<-rimp$neuron

r$R2<- r2

names(r)<- c(names(rimp$b[[1]]$lmg),'neuron','R2')
saveRDS(r,'allRelImpoSplit.RDS')
```
